Trajectories

### 2021-09-30 Rmd/manuscript.Rmd
Fix comprehension model and update report

documentclass     : "apa6"
classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine  : "pdflatex"

@@

knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  message = FALSE,
  warning = FALSE,
  cache.extra = knitr::rand_seed,
  out.width = "80%",
  results = "asis"
)
```

@@

We collected data from `r printnum(n_participants)` children from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`. All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX. Families were contacted by e-mail or phone if their child were aged between 10 and 34 months, and had not previously reported to be exposed more than 10% of the time to a language other than Spanish or Catalan. Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate. Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to the non-dominant languag (i.e., their degree of bilingualism).

```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Participant sample size by age and degree of exposure to a second language. Information about DoE was provided by families when filling the questionnaire. For illustration purposes this table, DoEs were binned into 10\\% wide bins, and ages (in months) were round to the nearest integer. Hyphens indicate that no participants from that specific combination of age filled the questionnaire."}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = age, values_from = n, values_fill = NA) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by age (months)", columns = 2:17) %>% 
  fmt_missing(everything(), everything(), missing_text = "-") %>% 
  cols_label(
    doe_2 = md("**L2 DoE**")
  ) %>% 
  tab_style(
    cell_text(style = "italic"),
    cells_column_labels(everything())
  ) %>% 
  tab_style(
    cell_text(weight = "bold"),
    cells_column_spanners(everything())
  )

@@

```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Educational attainment. Proportion of highest parental educational attainment by DoE to the second language. For illustration purposes, DoEs were binned into 10\\% wide bins."}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")

participants %>%
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  group_by(doe_2) %>%
  summarise(
    n_no_education = length(edu_parent[edu_parent=="No education"])/n(),
    n_primary = length(edu_parent[edu_parent=="Primary"])/n(),
    n_secondary = length(edu_parent[edu_parent=="Secondary"])/n(),
    n_complementary = length(edu_parent[edu_parent=="Complementary"])/n(),
    n_vocational = length(edu_parent[edu_parent=="Vocational"])/n(),
    n_university = length(edu_parent[edu_parent=="University"])/n(),
    .groups = "drop"
  ) %>% 
  gt() %>% 
  tab_spanner("Educational attainment", columns = 2:7) %>% 
  cols_label(
    doe_2 = md("**L2 DoE**"),
    n_no_education = "No education",
    n_primary = "Primary",
    n_secondary = "Secondary",
    n_complementary = "Complementary",
    n_vocational = "Vocational",
    n_university = "University"
  ) %>% 
  fmt_percent(2:7) %>% 
  fmt_missing(everything(), missing_text = "-") %>% 
  tab_style(
    cell_text(style = "italic"),
    cells_column_labels(everything())
  ) %>% 
  tab_style(
    cell_text(weight = "bold"),
    cells_column_spanners(everything())
  )
```

@@

```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Lexical frequencies. Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}
items %>%
  select(te, language, cognate, frequency) %>% 
  left_join(select(multilex_data$pool, te, language, version)) %>% 
  unnest(version) %>% 
  group_by(language, version, cognate) %>% 
  summarise(
    frequency_mean = mean(frequency, na.rm = TRUE),
    frequency_sd = sd(frequency, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>% 
  mutate(
    frequency_se = frequency_sd/sqrt(n),
    frequency_ci_lower = frequency_mean + qnorm(0.025)*frequency_se,
    frequency_ci_upper = frequency_mean + qnorm(0.975)*frequency_se
  ) %>% 
  pivot_wider(names_from = c("cognate"), values_from = c("n", matches("frequency"))) %>% 
  clean_names() %>% 
  mutate(
    frequency_ci_identical_cognate = paste0(
      "[", round(frequency_ci_lower_identical_cognate, 2), ", ", round(frequency_ci_lower_identical_cognate, 2), "]"
    ),
    frequency_ci_non_identical_cognate = paste0(
      "[", round(frequency_ci_lower_non_identical_cognate, 2), ", ", round(frequency_ci_lower_non_identical_cognate, 2), "]"
    ),
    frequency_ci_non_cognate = paste0(
      "[", round(frequency_ci_lower_non_cognate, 2), ", ", round(frequency_ci_lower_non_cognate, 2), "]"
    ),
  ) %>% 
  relocate(
    language, version,
    n_identical_cognate,
    frequency_mean_identical_cognate,
    frequency_se_identical_cognate,
    frequency_ci_identical_cognate,
    matches("_non_identical"),
    matches("_non_cognate")
  ) %>% 
  mutate(language = ifelse(language=="Catalan", "CAT", "SPA")) %>% 
  select(-matches("sd|upper|lower")) %>% 
  gt(groupname_col = "language") %>% 
  fmt_number(matches("mean|se")) %>% 
  tab_spanner("Identical cognates", columns = ends_with("_identical_cognate")) %>% 
  tab_spanner("Non-identical cognates", columns = ends_with("_non_identical_cognate")) %>% 
  tab_spanner("Non-cognates", columns = ends_with("_non_cognate")) %>% 
  cols_label(
    version = "List",
    n_identical_cognate = md("N"),
    n_non_identical_cognate = md("N"),
    n_non_cognate = md("N"),
    frequency_mean_identical_cognate = md("M"),
    frequency_mean_non_identical_cognate = md("M"),
    frequency_mean_non_cognate = md("M"),
    frequency_se_identical_cognate = md("SD"),
    frequency_se_non_identical_cognate = md("SD"),
    frequency_se_non_cognate = md("SD"),
    frequency_ci_identical_cognate = md("95% *CI"),
    frequency_ci_non_identical_cognate = md("95% CI"),
    frequency_ci_non_cognate = md("95% CI")
  ) %>% 
  cols_width(
    starts_with("frequency_ci_") ~ px(100),
    TRUE ~ px(40)
  ) %>% 
  tab_style(
    cell_text(style = "italic"),
    cells_column_labels(everything())
  ) %>% 
  tab_style(
    cell_text(weight = "bold"),
    cells_column_spanners(everything())
  )

@@

* *Age* of participant in months (`age`), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
* *Lexical frequency* of item (`frequency_center`), Relative lexical frequencies (frequency per million words, $fpmw$) were extracted from the CHILDES data base, using the `childesr` R package [@braginsky2018childesr], and were transformed to Zipf scores [@van2014subtlex; @zipf1949human] ($log10(fpmw)+3$). Missing frequencies were imputed by multiple imputation using the `mice` R package [van2011mice]. The resulting scores were then centred around the mean.
* *Degree of exposure* to the language (`doe_center`), computed as the proportion of the time the child is exposed to the language the item belongs to. For instance, for a response to the item `perro` (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1). Before entering the analyses, this variable was centred it around the mean and divided by 10, so that the values of the regression coefficients of this predictor correspond to a 10% increase in exposure to the language of the item.
* *Cognateness*, a categorical variable indicating whether the Catalan and Spanish forms of the TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist. We sum-coded two contrasts [@schad2020capitalize]: `cognate1`, comparing IC and NIC (both coded as +0.25) with NC (coded as -0.5) and `cognate2`, comparing IC (coded as +0.5) with NIC (coded as -0.5).

We first fit a base model that only included the main effects of `age` and `frequency_center`  (Model 1) [see @braginsky2019consistency; @jones2019children; @kachergis2021toward for similar approaches]. We then entered our predictors of interest by adding `doe_center` (Model 2) and then `cognate` and its interaction with `doe_center` (Model 3). To account for the hierarchical structure of the data (some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE )every model included *Participants* (`id`) and *TEs* (`te`) as grouping variables (i.e. random effects), with all possible random intercepts and slopes where appropriate [@barr2013random] (see Appendix for more details of the model).

@@

For both comprehension and production, the extended model--which included all main effects and interactions--fitted data the best: it showed the smallest absolute expected log predictive density. This estimate is several times larger than its standard error indicating high predictive accuracy. Table 3 summarises the LOO scores of the fitted models. In the next sections we report the fixed effects of the comprehension and production models separately. We report the mean of each coefficient's posterior distribution along with standard errors and 95% CrIs in Tables 4 and 5. For interpretability, we report the derivative of the logistic function for each coefficient^[The derivative is calculated as $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$. This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.] for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable [@gelman2020regression].

```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

l_comp <- loo_comp %>% 
  as_tibble() %>% 
  rownames_to_column("model") %>% 
  select(-matches("p_loo|looic")) %>% 
  relocate(model, matches("elpd_loo"), matches("diff")) %>%
  mutate_all(as.numeric) %>% 
  mutate(type = "Comprehension", model = paste0("Model ", model))

l_prod <- loo_prod %>% 
  as_tibble() %>% 
  rownames_to_column("model") %>% 
  select(-matches("p_loo|looic")) %>% 
  relocate(model, matches("elpd_loo"), matches("diff")) %>%
  mutate_all(as.numeric) %>% 
  mutate(type = "Production", model = paste0("Model ", model))

@@

gt(l, groupname_col = "type") %>% 
  tab_spanner(md("**ELPD</sub>**"), matches("elpd_loo")) %>% 
  tab_spanner(md("**Diff.**"), matches("diff")) %>%
  fmt_number(2:7) %>% 
  fmt_missing(columns = everything(), missing_text = "-") %>% 
  cols_label(
    model = "Model",
    # elpd
    elpd_loo = md("*ELPD*"),
    se_elpd_loo = md("*SE*"),
    subsampling_se_elpd_loo = md("*SE* (sub)"),
    # diff
    elpd_diff = md("*diff*"),
    se_diff = md("*SE*"),
    subsampling_se_diff = md("*SE* (sub)"),
  )
```

# Comprehension

The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_comp["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0: mean participant age (`r printnum(mean(responses$age))` months), mean item lexical frequency (`r printnum(mean(responses$frequency))`, Zipf score), mean participant DoE (`r printnum(mean(responses$doe_center))`%), and regardless of its cognate status (average acros the three levels). An increase in one standard deviation in age (1 *SD* = `r printnum(sd(responses$age))` months) was associated with a `r printnum(coef_comp["age", "Estimate"]*100/4)`% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_comp["frequency", "Estimate"]*100/4)`% increment in comprehension probability. A 10% increase in relative language exposure made TEs `r printnum(coef_comp["doe_center", "Estimate"]*100/4)`% more likely to be understood. Cognates (identical and non-identical) were `r printnum(coef_comp["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates. The 95% credible interval of the `doe_center:cognate1` interaction did not exclude 0. Identical cognates were `r printnum(abs(coef_comp["cognate2", "Estimate"]*100/4))`% less like to be understood than non-identical cognates. The 95% credible interval of the `doe_center:cognate2` interaction excluded 0 (although its lower bound was near zero). Figure 1 summarises the posterior distribution of the fixed effects of the model.

```{r fix_effets_comp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for comprehension. Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

c_comp <- fixef(fits_comp$fit_2) %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  clean_names() %>% 
  mutate(
    estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
    q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
    q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
    est_error = ifelse(term=="Intercept", inv_logit_scaled(est_error), est_error/4)
  ) %>% 
  rename_at(vars(-matches("term")), ~paste0("comp_", .)) %>% 
  select(-one_of(c("rhat", "bulk_ess", "tail_ess")))

@@

gt(c) %>% 
  fmt_percent(matches("estimate|error")) %>% 
  fmt_percent(matches("_q")) %>% 
  tab_spanner("Comprehension", matches("comp_")) %>% 
  tab_spanner("Production", matches("prod_")) %>% 
  cols_merge(vars("comp_q2_5", "comp_q97_5"), pattern = "[{1}, {2}]") %>% 
  cols_merge(vars("prod_q2_5", "prod_q97_5"), pattern = "[{1}, {2}]") %>%
  cols_label(
    term = md("**Predictor**"),
    comp_estimate = md("**Mean**"),
    comp_est_error = md("**SE**"),
    comp_q2_5 = md("**95\\% CrI**"),
    comp_q2_5 = md("**95\\% CrI**"),
    prod_q2_5 = md("**95\\% CrI**"),
    prod_estimate = md("**Mean**"),
    prod_est_error = md("**SE**"),
    prod_q2_5 = md("**95\\% CrI**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
  ) %>% 
  cols_align(
    align = c("left"),
    columns = 2:4
  ) 
```

>>>>

documentclass     : "apa6"
classoption       : "man"
output            :
  word_document:
    reference_docx: "manuscript_template.docx"

@@

knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  message = FALSE,
  warning = FALSE,
  cache.extra = knitr::rand_seed,
  out.width = "100%",
  dpi = 300,
  results = "asis"
)
```

@@

We collected data from `r printnum(n_participants)` children from the Metropolitan Area of Barcelona between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`. All families gave informed consent before participating and this study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX. Families were contacted by e-mail or phone if their child were aged between 10 and 34 months, and had not previously reported to be exposed more than 10% of the time to a language other than Spanish or Catalan. Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device in a browser within the two week following the invitation to participate. Table 1 summarises the distribution of participants across ages and degrees of exposure (DoE) to the non-dominant language (i.e., their degree of bilingualism).

```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Participant sample size by age and degree of exposure to a second language. Information about DoE was provided by families when filling the questionnaire. For illustration purposes this table, DoEs were binned into 10\\% wide bins, and ages (in months) were round to the nearest integer. Hyphens indicate that no participants from that specific combination of age filled the questionnaire."}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
t <- participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = age, values_from = n, values_fill = NA) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by age (months)", columns = 2:17) %>% 
  fmt_missing(everything(), everything(), missing_text = "-") %>% 
  cols_label(
    doe_2 = md("**L2 DoE**")
  ) %>% 
  tab_style(
    cell_text(style = "italic"),
    cells_column_labels(everything())
  ) %>% 
  tab_style(
    cell_text(weight = "bold"),
    cells_column_spanners(everything())
  )

gtsave(t, here("Figures/participants_lp.png"))

@@

```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Educational attainment. Proportion of highest parental educational attainment by DoE to the second language. For illustration purposes, DoEs were binned into 10\\% wide bins."}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")

t <- participants %>%
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  group_by(doe_2) %>%
  summarise(
    n_no_education = length(edu_parent[edu_parent=="No education"])/n(),
    n_primary = length(edu_parent[edu_parent=="Primary"])/n(),
    n_secondary = length(edu_parent[edu_parent=="Secondary"])/n(),
    n_complementary = length(edu_parent[edu_parent=="Complementary"])/n(),
    n_vocational = length(edu_parent[edu_parent=="Vocational"])/n(),
    n_university = length(edu_parent[edu_parent=="University"])/n(),
    .groups = "drop"
  ) %>% 
  gt() %>% 
  tab_spanner("Educational attainment", columns = 2:7) %>% 
  cols_label(
    doe_2 = md("**L2 DoE**"),
    n_no_education = "No education",
    n_primary = "Primary",
    n_secondary = "Secondary",
    n_complementary = "Complementary",
    n_vocational = "Vocational",
    n_university = "University"
  ) %>% 
  fmt_percent(2:7) %>% 
  fmt_missing(everything(), missing_text = "-") %>% 
  tab_style(
    cell_text(style = "italic"),
    cells_column_labels(everything())
  ) %>% 
  tab_style(
    cell_text(weight = "bold"),
    cells_column_spanners(everything())
  )

gtsave(t, here("Figures/participants_edu.png"))

@@

```{r items_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies. Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}

d <- items %>%
  select(te, language, cognate, frequency) %>% 
  left_join(select(multilex_data$pool, te, language, version)) %>% 
  unnest(version) %>% 
  group_by(language, version, cognate) %>% 
  summarise(
    frequency_mean = mean(frequency, na.rm = TRUE),
    frequency_sd = sd(frequency, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>% 
  mutate(
    frequency_se = frequency_sd/sqrt(n),
    frequency_ci_lower = frequency_mean + qnorm(0.025)*frequency_se,
    frequency_ci_upper = frequency_mean + qnorm(0.975)*frequency_se
  ) %>% 
  pivot_wider(names_from = c("cognate"), values_from = c("n", matches("frequency"))) %>% 
  clean_names() %>% 
  mutate(
    frequency_ci_identical_cognate = paste0(
      "[", round(frequency_ci_lower_identical_cognate, 2), ", ", round(frequency_ci_lower_identical_cognate, 2), "]"
    ),
    frequency_ci_non_identical_cognate = paste0(
      "[", round(frequency_ci_lower_non_identical_cognate, 2), ", ", round(frequency_ci_lower_non_identical_cognate, 2), "]"
    ),
    frequency_ci_non_cognate = paste0(
      "[", round(frequency_ci_lower_non_cognate, 2), ", ", round(frequency_ci_lower_non_cognate, 2), "]"
    ),
  ) %>% 
  relocate(
    language, version,
    n_identical_cognate,
    frequency_mean_identical_cognate,
    frequency_se_identical_cognate,
    frequency_ci_identical_cognate,
    matches("_non_identical"),
    matches("_non_cognate")
  ) %>% 
  mutate(language = ifelse(language=="Catalan", "CAT", "SPA")) %>% 
  select(-matches("sd|upper|lower"))

t <- d %>% 
  gt(groupname_col = "language") %>% 
  fmt_number(matches("mean|se")) %>% 
  tab_spanner("Identical cognates", columns = ends_with("_identical_cognate")) %>% 
  tab_spanner("Non-identical cognates", columns = ends_with("_non_identical_cognate")) %>% 
  tab_spanner("Non-cognates", columns = ends_with("_non_cognate")) %>% 
  cols_label(
    version = "List",
    n_identical_cognate = md("N"),
    n_non_identical_cognate = md("N"),
    n_non_cognate = md("N"),
    frequency_mean_identical_cognate = md("M"),
    frequency_mean_non_identical_cognate = md("M"),
    frequency_mean_non_cognate = md("M"),
    frequency_se_identical_cognate = md("SD"),
    frequency_se_non_identical_cognate = md("SD"),
    frequency_se_non_cognate = md("SD"),
    frequency_ci_identical_cognate = md("95% *CI"),
    frequency_ci_non_identical_cognate = md("95% CI"),
    frequency_ci_non_cognate = md("95% CI")
  ) %>% 
  cols_width(
    starts_with("frequency_ci_") ~ px(100),
    TRUE ~ px(40)
  ) %>% 
  tab_style(
    cell_text(style = "italic"),
    cells_column_labels(everything())
  ) %>% 
  tab_style(
    cell_text(weight = "bold"),
    cells_column_spanners(everything())
  )

gtsave(t, here("Figures/item_frequency.png"))

@@

* *Age* of participant in months (`age_center`), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
* *Lexical frequency* of item (`frequency_center`), Relative lexical frequencies (frequency per million words, $fpmw$) were extracted from the CHILDES data base, using the `childesr` R package [@braginsky2018childesr], and were transformed to Zipf scores [@van2014subtlex; @zipf1949human] ($log10(fpmw)+3$). Missing frequencies were imputed by multiple imputation using the `mice` R package [van2011mice]. The resulting scores were then centred around the mean.
* *Degree of exposure* to the language (`doe_center`), computed as the proportion of the time the child is exposed to the language the item belongs to. For instance, for a response to the item `perro` (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1). Before entering the analyses, this variable was centred it around the mean and divided by 10, so that the values of the regression coefficients of this predictor correspond to a 10% increase in exposure to the language of the item.
* *Cognateness*, a categorical variable indicating whether the Catalan and Spanish forms of the TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist. We sum-coded two contrasts [@schad2020capitalize]: `cognate1`, comparing IC and NIC (both coded as +0.25) with NC (coded as -0.5) and `cognate2`, comparing IC (coded as +0.5) with NIC (coded as -0.5).

We first fit a base model that only included the main effects of `age_center` and `frequency_center`  (Model 1) [see @braginsky2019consistency; @jones2019children; @kachergis2021toward for similar approaches]. We then entered our predictors of interest by adding `doe_center` (Model 2) and then `cognate` and its interaction with `doe_center` (Model 3). To account for the hierarchical structure of the data (some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE )every model included *Participants* (`id`) and *TEs* (`te`) as grouping variables (i.e. random effects), with all possible random intercepts and slopes where appropriate [@barr2013random] (see Appendix for more details of the model).

@@

For both comprehension and production, the extended model--which included all main effects and interactions--fitted data the best: it showed the smallest absolute expected log predictive density. This estimate is several times larger than its standard error indicating high predictive accuracy. Table 3 summarises the LOO scores of the fitted models. In the next sections we report the fixed effects of the comprehension and production models separately. We report the mean of each coefficient's posterior distribution along with standard errors and 95% credible intervals (CrI) in Tables 4 and 5. For interpretability, we report the derivative of the logistic function for each coefficient^[The derivative is calculated as $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$. This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.] for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable [@gelman2020regression].

```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Leave-one-out cross-validation. The expected log predictive density (*ELPD*) indicates the sum of the likelihood of each observation, given the model when fitted without such obervation. Values close to 0 indicate that the model fits data better. Standard errors (*SE*) indicate the uncertainty around the computed ELPD in each model. A small *SE* (in realtion to the magnitde of the *ELPD* indicates less uncertainty around the ELPD estimate."}

l_comp <- loo_comp %>% 
  as_tibble() %>% 
  select(-matches("p_loo|looic")) %>% 
  mutate_all(as.numeric) %>% 
  mutate(
    type = "Comprehension",
    model = paste0("Model ", 2:0)
  ) %>% 
  relocate(model, type, matches("elpd_loo"), matches("diff"))

l_prod <- loo_prod %>% 
  as_tibble() %>% 
  select(-matches("p_loo|looic")) %>% 
  mutate_all(as.numeric) %>% 
  mutate(
    type = "Production",
    model = paste0("Model ", 2:0)
  ) %>% 
  relocate(model, type, matches("elpd_loo"), matches("diff"))

@@

t <- gt(l, groupname_col = "type") %>% 
  tab_spanner(md("**ELPD**"), matches("elpd_loo")) %>% 
  tab_spanner(md("**Diff.**"), matches("diff")) %>%
  fmt_number(2:7) %>% 
  fmt_missing(columns = everything(), missing_text = "-") %>% 
  cols_label(
    model = "Model",
    # elpd
    elpd_loo = md("*ELPD*"),
    se_elpd_loo = md("*SE*"),
    subsampling_se_elpd_loo = md("*SE* (sub)"),
    # diff
    elpd_diff = md("*diff*"),
    se_diff = md("*SE*"),
    subsampling_se_diff = md("*SE* (sub)"),
  )

gtsave(t, here("Figures/loos.png"))

## Comprehension

The average probability of participants understanding a TE was `r printnum(inv_logit_scaled(coef_comp["Intercept", "Estimate"])*100)`% (intercept) when all predictors were equal to 0: mean participant age (`r printnum(mean(responses$age))` months), mean item lexical frequency (`r printnum(mean(responses$frequency))`, Zipf score), mean participant DoE (`r printnum(mean(responses$doe)*10)`%), and regardless of its cognate status (average across the three levels). An increase in one month of age was associated with a `r printnum(coef_comp["age", "Estimate"]*100/4)`% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency))` Zipf scores) was associated with a `r printnum(coef_comp["frequency", "Estimate"]*100/4)`% increment in comprehension probability. A 10% increase in relative language exposure made TEs `r printnum(abs(coef_comp["doe_center", "Estimate"]*100/4))`% less likely to be understood. Cognates (identical and non-identical) were `r printnum(coef_comp["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates. The 95% credible interval of the `doe_center:cognate1` interaction excluded 0 and covered negative values, indicating that for every 10% increment in DoE, the difference in comprehension probability between cognates and non-cognates decreased in `r printnum(abs(coef_comp["doe_center:cognate1", "Estimate"]*100/4))`%. Identical cognates were `r printnum(abs(coef_comp["cognate2", "Estimate"]*100/4))`% less like to be understood than non-identical cognates. The 95% credible interval of the `doe_center:cognate2` interaction excluded 0 (although its lower bound was near zero), and covered only negative values, indicating that for every 10% increase in DoE, the difference in comprehension probability between identical cognates and non-identical cognates decreased in `r printnum(abs(coef_comp["doe_center:cognate2", "Estimate"]*100/4))`%. Figure 1 summarises the posterior distribution of the fixed effects of the model.

```{r fix_effets, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Point estimate (mean) and credible interval (95\\%) of the posterior distribution of the coefficients of the fixed effects of the best fitting model for comprehension and production. Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The rest of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

c_comp <- fixef(fits_comp$fit_2) %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  clean_names() %>% 
  mutate(
    estimate = ifelse(term=="Intercept", inv_logit_scaled(estimate), estimate/4),
    q2_5 = ifelse(term=="Intercept", inv_logit_scaled(q2_5), q2_5/4),
    q97_5 = ifelse(term=="Intercept", inv_logit_scaled(q97_5), q97_5/4),
    est_error = ifelse(term=="Intercept", inv_logit_scaled(est_error), est_error/4)
  ) %>% 
  rename_at(vars(-matches("term")), ~paste0("comp_", .)) %>% 
  select(-any_of(c("rhat", "bulk_ess", "tail_ess")))

@@

t <- gt(c) %>% 
  fmt_percent(matches("estimate|error")) %>% 
  fmt_percent(matches("_q")) %>% 
  tab_spanner("Comprehension", matches("comp_")) %>% 
  tab_spanner("Production", matches("prod_")) %>% 
  cols_merge(vars("comp_q2_5", "comp_q97_5"), pattern = "[{1}, {2}]") %>% 
  cols_merge(vars("prod_q2_5", "prod_q97_5"), pattern = "[{1}, {2}]") %>%
  cols_label(
    term = md("**Predictor**"),
    comp_estimate = md("**Mean**"),
    comp_est_error = md("**SE**"),
    comp_q2_5 = md("**95\\% CrI**"),
    comp_q2_5 = md("**95\\% CrI**"),
    prod_q2_5 = md("**95\\% CrI**"),
    prod_estimate = md("**Mean**"),
    prod_est_error = md("**SE**"),
    prod_q2_5 = md("**95\\% CrI**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age_center", "frequency_center", "doe_center", "cognate1", "cognate2", "doe_center:cognate1"))
  ) %>% 
  cols_align(
    align = c("left"),
    columns = 2:4
  )
gtsave(t, here("Figures/fix_effects.png"))
